---
title: "MSOA walking distances"
output:
  html_document:
    df_print: paged
---
```{r}
knitr::opts_chunk$set(eval = FALSE
                      ,include = TRUE
                      ,echo = TRUE)
#library(osmextract)
library(lwgeom)
library(dodgr)
library(osmdata)
library(pct)
library(dplyr)
library(ggplot2)
library(sf)
library(tmap)
library(tidyverse)
library(units)
library(reshape2)
```

1. LOADING DATA: 

1.1 Download MSOA boundaries
```{r}
#Polygons:
london_msoa = get_pct_zones(region = "london", geography = "msoa")
```

 1.2 Download OS Open Road data:
```{r}
#getwd()
London_OS <- sf::read_sf("oproad_essh_gb/data/TQ_RoadLink.shp")
London_OS <- st_transform(London_OS, st_crs(london_msoa))
bbox_London<- st_bbox(london_msoa)

#Spatial subsetting using MSOA bounding box:
London_OS <- London_OS[st_as_sfc(bbox_London), ]  
```
 1.3 Load OSM roads:
```{r}
# From England dataset, not greater london because of boundary effect

#osm_roads <- sf::read_sf("england-latest-free/gis_osm_roads_free_1.shp")
#osm_roads <- st_transform(osm_roads, st_crs(london_msoa))
#London_OSM <- osm_roads[st_as_sfc(bbox_London), ] 
#st_write(London_OSM, getwd(), "London_OSM.shp",driver = "ESRI Shapefile")

London_OSM <- sf::read_sf("London_OSM.shp")
London_OSM <- st_transform(London_OSM, st_crs(london_msoa))

```

```{r}
#unique(London_OSM$fclass)
London_OSM=London_OSM %>% dplyr::filter(!fclass %in% c("motorway_link", "motorway"))
```


2. OD POINTS:

2.1 MSOA centroids (nearest to OSM residential streets)

```{r}
#get MSOA centroids
msoa_centroid_geom <- london_msoa %>% 
  st_centroid() 

# Centroids nearest to OSM net points
London_OD=London_OSM %>% dplyr::filter(fclass %in% c("residential", "living_street"))

London_OD_ptns = st_cast(London_OD, "POINT")

nearest=st_nearest_feature(msoa_centroid_geom, London_OD_ptns)
london_centroid_geom=London_OD_ptns[nearest,]
london_centroid_geom$geo_code=msoa_centroid_geom$geo_code
london_centroid_geom=london_centroid_geom %>% select("geo_code", "geometry","osm_id")


#centroids geo codes 
geo_code=data.frame(geo=london_centroid_geom$geo_code,idgeo=(1:length(london_centroid_geom$geo_code)))

# get coords:
london_centroid_coords=london_centroid_geom%>% 
    st_coordinates()


st_write(london_centroid_geom, "C:/Users/vales/OneDrive/Desktop/Pedestrian network/data", "london_centroid_geom.shp",driver = "ESRI Shapefile",append=FALSE)
write.csv(london_centroid_coords,"C:/Users/vales/OneDrive/Desktop/Pedestrian network/data/london_centroid_coords.csv", row.names = FALSE)
```


2.2 Sample of nodes (from OS layer)
```{r}
#edges within msoa
london_OSM_inMsoa=st_join(London_OD, london_msoa) %>% select("osm_id", "fclass", "geometry", "geo_code")

# points within edges
nodes_london_OSM_inMsoa <- london_OSM_inMsoa %>%
  st_coordinates() %>%
  as_tibble() 

# as WKT geometry column
nodes_london_OSM_inMsoa=nodes_london_OSM_inMsoa%>% 
  st_as_sf(coords = c("X", "Y"))

st_crs(nodes_london_OSM_inMsoa) <- st_crs(london_msoa) 

# select random %
sample_node=nodes_london_OSM_inMsoa[sample(length(nodes_london_OSM_inMsoa$geometry),(length(nodes_london_OSM_inMsoa$geometry)*0.03)),]

# Nodes within msoa boundaries
sample_node_within=sample_node[london_msoa,]

sample_node_within=st_join(sample_node_within, london_msoa) %>% select("L1", "geometry", "geo_code")


sample_node_coords <- sample_node_within %>%
  mutate(lat = unlist(map(sample_node_within$geometry,1)),
         long = unlist(map(sample_node_within$geometry,2)))

sample_node_coords=as.data.frame(sample_node_coords)
sample_node_coords=sample_node_coords[,c(3,4,5)]
colnames(sample_node_coords)=c("geo_code","X","Y")
sample_node_coords$id  <- 1:nrow(sample_node_coords)

write.csv(sample_node_coords,"data/sample_nodes_osm.csv", row.names = FALSE)

st_write(sample_node_within, "C:/Users/vales/OneDrive/Desktop/Pedestrian network/data", "sample_node_within_geom.shp",driver = "ESRI Shapefile",append=FALSE)

```


```{r}
tmap_mode("view")
qtm(london_msoa, fill=NULL, borders = "red")+
  qtm(sample_node_within, symbols.col = "red", symbols.size=0.01)

```
3. EUCLIDIAN DISTANCES


3.0 Reads OD points 
```{r}

#centroids:
london_centroid_coords <- read.csv(file = 'data/london_centroid_coords.csv')
london_centroid_geom <- sf::read_sf("data/london_centroid_geom.shp")

#sample nodes:
sample_node_coords <- read.csv(file = 'data/sample_nodes_osm.csv')
sample_node_within <- sf::read_sf("data/sample_node_within_geom.shp")

#geo_code:
geo_code=data.frame(geo=london_centroid_geom$geo_code,idgeo=(1:length(london_centroid_geom$geo_code)))
```


3.1 MSOA centroids:

```{r}
Centroid_eu <- london_centroid_geom %>% 
  st_distance() %>%
  set_units(NULL)

# To dataframe
Centroid_eu_df=setNames(melt(Centroid_eu), c('rows', 'cols', 'cents_eu'))

Centroid_eu_df <- merge(Centroid_eu_df,geo_code, by.x = "rows",by.y = "idgeo")
Centroid_eu_df <- merge(Centroid_eu_df,geo_code, by.x = "cols",by.y = "idgeo")
colnames(Centroid_eu_df)<-c( "cols", "rows", "cents_eu","geo_code_r","geo_code_c")
```

3.2 Sample nodes:
```{r}
sample_node_within=sample_node_within[,2:3]
head(sample_node_within)

Nodes_eu <- sample_node_within %>% 
  st_distance() %>%
  set_units(NULL)

colnames(Nodes_eu) <- c(sample_node_within$geo_code)
rownames(Nodes_eu) <- c(sample_node_within$geo_code)



# To dataframe
Nodes_eu_df=setNames(melt(Nodes_eu), c('geo_code_r', 'geo_code_c', 'nodes_eu'))

Nodes_eu_df_mean=Nodes_eu_df %>%
  group_by(geo_code_r , geo_code_c) %>%
  summarise(nodes_eu_mean=(mean(nodes_eu)),.groups = 'drop')

head(Nodes_eu_df_mean)

```




```{r}
euclidian=merge(Nodes_eu_df_mean, Centroid_eu_df, by=c("geo_code_r","geo_code_c"))
euclidian=euclidian %>% select("geo_code_r","geo_code_c","cents_eu","nodes_eu_mean")

write.csv(euclidian,"data/euclidian.csv", row.names = FALSE)
```


4. COMPUTE NETWORK DISTANCES

4.1 Network distances between centroids

4.1.1 OS net between centroids:

```{r}
# Convert OS to 'dodgr' network:

# specification of weighting profile:
colnm <- "formOfWay" # name of column used to determine weights
wts <- data.frame (name = "custom",
                   way = unique (London_OS [[colnm]]),
                   value = c (1, 1, 1, 1,1,1))

London_OS_net <- weight_streetnet (London_OS, wt_profile = wts,
                                   type_col = colnm, id_col = "identifier")
# 'id_col' tells the function which column to use to attribute IDs of ways

London_OS_net <- London_OS_net[London_OS_net$component == 1, ]

```

```{r}
# Compute distances:

centroid_net_OS <- dodgr_dists(London_OS_net
                               ,from = london_centroid_coords
                               ,to = london_centroid_coords
                               ,shortest = TRUE
                               ,quiet = FALSE
                               ,parallel=TRUE)



# To dataframe
Centroid_net_OS_df=setNames(melt(centroid_net_OS), c('rows', 'cols', 'cents_os'))
Centroid_net_OS_df$id  <- 1:nrow(Centroid_net_OS_df)

Centroid_net_OS_df <- merge(Centroid_net_OS_df,geo_code, by.x = "rows",by.y = "idgeo")
Centroid_net_OS_df <- merge(Centroid_net_OS_df,geo_code, by.x = "cols",by.y = "idgeo")
colnames(Centroid_net_OS_df)<-c( "cols", "rows", "cents_os","id" ,"geo_code_r","geo_code_c")

```


4.1.2 OSM net between centroids:
```{r}
#Create dodgr network

# specification of weighting profile:
colnm <- "fclass" # name of column used to determine weights
wts <- data.frame (name = "custom",
                   way = unique (London_OSM [[colnm]]),
                   value = c (rep(1, each = length(unique(London_OSM$fclass)))))

London_OSM_net <- weight_streetnet (London_OSM, wt_profile = wts,
                                   type_col = colnm, id_col = "identifier")
# 'id_col' tells the function which column to use to attribute IDs of ways


London_OSM_net <- London_OSM_net[London_OSM_net$component == 1, ]
```


```{r}

#Compute distances: 
centroid_net_OSM <- dodgr_dists(London_OSM_net
                               ,from = london_centroid_coords
                               ,to = london_centroid_coords
                               ,shortest = TRUE
                               ,quiet = FALSE
                               ,parallel=TRUE)


# To dataframe
Centroid_net_OSM_df=setNames(melt(centroid_net_OSM), c('rows', 'cols', 'cents_osm'))
Centroid_net_OSM_df$id  <- 1:nrow(Centroid_net_OSM_df)

Centroid_net_OSM_df <- merge(Centroid_net_OSM_df,geo_code, by.x = "rows",by.y = "idgeo")
Centroid_net_OSM_df <- merge(Centroid_net_OSM_df,geo_code, by.x = "cols",by.y = "idgeo")
colnames(Centroid_net_OSM_df)<-c( "cols", "rows", "cents_osm","id" ,"geo_code_r","geo_code_c")

```

```{r}
Centroids_net=merge(Centroid_net_OS_df, Centroid_net_OSM_df, by=c("geo_code_r","geo_code_c"))
Centroids_net=Centroids_net %>% select("geo_code_r","geo_code_c","cents_os","cents_osm")

write.csv(Centroids_net,"data/Centroids_net.csv", row.names = FALSE)
```


4.2 Network distances between sample nodes

4.2.1 OS net between sample nodes:
```{r}
gc()
memory.limit(size=30000)

nodes_net_OS <- dodgr_dists(London_OS_net
                            ,from = sample_node_coords
                            ,to = sample_node_coords
                            ,shortest = TRUE
                            ,quiet = FALSE
                            ,parallel=TRUE)



colnames(nodes_net_OS) <- c(sample_node_coords$geo_code)
rownames(nodes_net_OS) <- c(sample_node_coords$geo_code)



# Nodes distance from matrix to dataframe:
Nodes_net_OS_df=setNames(melt(nodes_net_OS), c('geo_code_r', 'geo_code_c', 'nodes_os'))

head(Nodes_net_OS_df)



Nodes_net_OS_df_mean=Nodes_net_OS_df %>%
  group_by(geo_code_r , geo_code_c) %>%
  summarise(nodes_os_mean=(mean(nodes_os)),.groups = 'drop')

head(Nodes_net_OS_df_mean)


write.csv(Nodes_net_OS_df_mean,"data/Nodes_net_OS_df_mean.csv", row.names = FALSE)
```


4.2.2 OSM net between sample nodes:
```{r}
gc()
memory.limit(size=50000)

nodes_net_OSM <- dodgr_dists(London_OSM_net
                            ,from = sample_node_coords
                            ,to = sample_node_coords
                            ,shortest = TRUE
                            ,quiet = FALSE
                            ,parallel=TRUE)

colnames(nodes_net_OSM) <- c(sample_node_coords$geo_code)
rownames(nodes_net_OSM) <- c(sample_node_coords$geo_code)


# Nodes distance from matrix to dataframe:
Nodes_net_OSM_df=setNames(melt(nodes_net_OSM), c('geo_code_r', 'geo_code_c', 'nodes_osm'))




#Average distances between nodes
Nodes_net_OSM_df_mean=Nodes_net_OSM_df %>%
  group_by(geo_code_r , geo_code_c) %>%
  summarise(nodes_osm_mean=(mean(nodes_osm)),.groups = 'drop')

head(Nodes_net_OSM_df_mean)




write.csv(Nodes_net_OSM_df_mean,"data/Nodes_net_OSM_df_mean.csv", row.names = FALSE)
```

```{r}

#merge sample nodes_net distances:
nodes_net=merge(Nodes_net_OSM_df_mean, Nodes_net_OS_df_mean, by=c("geo_code_r","geo_code_c"))

#merge net distances
all_net=merge(Centroids_net, nodes_net, by=c("geo_code_r","geo_code_c"))


#merge all distances
all_distances=merge(all_net, euclidian, by=c("geo_code_r","geo_code_c"))

head(all_distances)

write.csv(all_distances,"data/all_distances.csv", row.names = FALSE)

```


5. WEIGHTED CENTROIDS
```{r}
library(data.table)
library(jsonlite)

msoa_wcent <- read_json("msoa_pop_weighted.geojson"
                  ,simplifyVector = TRUE)

msoa_wcent <- msoa_wcent$features

msoa_wcent_geom <- msoa_wcent$geometry$coordinates %>%
  lapply(st_point) %>%
  st_sfc(crs = 4326) %>%
  st_sf()

msoa_wcent_geom$geo_code=msoa_wcent$properties$msoa11cd



# Centroids nearest to OSM net points
London_OD=London_OSM %>% dplyr::filter(fclass %in% c("residential", "living_street"))

London_OD_ptns = st_cast(London_OD, "POINT")
nearest=st_nearest_feature(msoa_wcent_geom, London_OD_ptns)
w_centroid_geom=London_OD_ptns[nearest,]
w_centroid_geom$geo_code=msoa_wcent_geom$geo_code
w_centroid_geom=w_centroid_geom %>% select("geo_code", "geometry","osm_id")

#centroids geo codes 
geo_code=data.frame(geo=w_centroid_geom$geo_code,idgeo=(1:length(w_centroid_geom$geo_code)))

# get coords:
w_centroid_coords=w_centroid_geom%>% 
    st_coordinates()


st_write(w_centroid_geom, "C:/Users/vales/OneDrive/Desktop/Pedestrian network/data", "w_centroid_geom.shp",driver = "ESRI Shapefile",append=FALSE)
write.csv(w_centroid_coords,"C:/Users/vales/OneDrive/Desktop/Pedestrian network/data/w_centroid_coords.csv", row.names = FALSE)
```
5.1 Euclidian distances:
```{r}
w_centroid_eu <- w_centroid_geom %>% 
  st_distance() %>%
  set_units(NULL)

# To dataframe
w_centroid_eu_df=setNames(melt(w_centroid_eu), c('rows', 'cols', 'w_cents_eu'))

w_centroid_eu_df <- merge(w_centroid_eu_df,geo_code, by.x = "rows",by.y = "idgeo")
w_centroid_eu_df <- merge(w_centroid_eu_df,geo_code, by.x = "cols",by.y = "idgeo")
colnames(w_centroid_eu_df)<-c( "cols", "rows", "w_cents_eu","geo_code_r","geo_code_c")
```
5.2 Net distances:
```{r}

w_centroid_net_OS <- dodgr_dists(London_OS_net
                               ,from = w_centroid_coords
                               ,to = w_centroid_coords
                               ,shortest = TRUE
                               ,quiet = FALSE
                               ,parallel=TRUE)



# To dataframe
w_centroid_net_OS_df=setNames(melt(w_centroid_net_OS), c('rows', 'cols', 'w_cents_os'))
w_centroid_net_OS_df$id  <- 1:nrow(w_centroid_net_OS_df)

w_centroid_net_OS_df <- merge(w_centroid_net_OS_df,geo_code, by.x = "rows",by.y = "idgeo")
w_centroid_net_OS_df <- merge(w_centroid_net_OS_df,geo_code, by.x = "cols",by.y = "idgeo")
colnames(w_centroid_net_OS_df)<-c( "cols", "rows", "w_cents_os","id" ,"geo_code_r","geo_code_c")
```

```{r}
#Compute distances: 
w_centroid_net_OSM <- dodgr_dists(London_OSM_net
                               ,from = w_centroid_coords
                               ,to = w_centroid_coords
                               ,shortest = TRUE
                               ,quiet = FALSE
                               ,parallel=TRUE)


# To dataframe
w_centroid_net_OSM_df=setNames(melt(w_centroid_net_OSM), c('rows', 'cols', 'w_cents_osm'))
w_centroid_net_OSM_df$id  <- 1:nrow(w_centroid_net_OSM_df)

w_centroid_net_OSM_df <- merge(w_centroid_net_OSM_df,geo_code, by.x = "rows",by.y = "idgeo")
w_centroid_net_OSM_df <- merge(w_centroid_net_OSM_df,geo_code, by.x = "cols",by.y = "idgeo")
colnames(w_centroid_net_OSM_df)<-c( "cols", "rows", "w_cents_osm","id" ,"geo_code_r","geo_code_c")
```

```{r}
w_cent_net=merge(w_centroid_net_OSM_df, w_centroid_net_OS_df, by=c("geo_code_r","geo_code_c"))
w_cent_net=w_cent_net %>% select("geo_code_r","geo_code_c","w_cents_osm","w_cents_os")

write.csv(w_cent_net,"data/w_cent_net.csv", row.names = FALSE)
```

```{r}
w_cents_all=merge(w_cent_net, w_centroid_eu_df, by=c("geo_code_r","geo_code_c"))
w_cents_all=w_cents_all %>% select("geo_code_r","geo_code_c","w_cents_osm","w_cents_os","w_cents_eu")
```


```{r}
all_distances=merge(all_distances, w_cents_all, by=c("geo_code_r","geo_code_c"))
```

```{r}
write.csv(all_distances,"data/all_distances.csv", row.names = FALSE)
```

